---
title: "Scanfish from OpenEarth OPeNDAP"
author: "Willem Stolte"
date: "January 8, 2016"
output: html_document
---

```{r load_packages, echo = F, message=F, warning=F, results=F, error=F}
require(ncdf4)
require(rworldmap)
library(ggplot2)
library(scales)
library(plyr)
```

### Download netcdf scanfish data

Scanfish data are available via the OpenEarth OPeNDAP server at Deltares. The NetCDF file is first downloaded to a local directory before it is read into memory. Direct acces to NetCDF files using R is possible but this is not yet a standard procedure under MS-Windows. 

```{r dowload_and_read_NetCDF, warning=F, comment= F}

url_scanfish <-"http://opendap.deltares.nl/thredds/fileServer/opendap/rijkswaterstaat/scanfish/scanfish_temperature.nc" 

# download the netcdf file
download.file(url_scanfish, "scanfish_temperature.nc" , method = "auto",
              quiet = FALSE, mode="wb", cacheOK = TRUE)

# check directory
list.files(pattern = ".nc")

# open connection to nc file
con <- nc_open("scanfish_temperature.nc")

# print(con)  # to see what's inside
```

Before plotting, the data are converted into a dataframe. the RNetCDF package function read.nc is used for this in combination with the standard function data.frame. 

```{r make_dataframe, warning=F, comment= F}
# read data into dataframe
z <- ncvar_get(con, var = "z")
lon <- ncvar_get(con, var = "lon")
lat <- ncvar_get(con, var = "lat")
temp <- ncvar_get(con, var = "temperature")
time <- ncvar_get(con, var = "TIME")
df <- data.frame(cbind(z, lon, lat, time, temp))


# turn time array into something useful
df$date <- as.Date(df$time, origin = "1970-01-01 00:00:00", tz = "UTC")
df$datetime <- as.POSIXct(df$date)
df$months <- format(df$date, "%m")
```

The data are now plotted on a static map, using a background map from the rworldmap package. 

```{r plot, warning=F, comment= F, fig.width= 7, fig.height= 6}
# load country polygon data
data(countriesLow)
# turn into dataframe for plotting
world <- fortify(countriesLow)
# make subset of scanfish data
df2 <- subset(df, df$month %in% c("07", "08", "09"))
# define bbox for plotting
xxlim <- expand_range(range(df2$lon), mul = 0.1)
yylim <- expand_range(range(df2$lat), mul = 0.1)

p <- ggplot(df2, aes(lon, lat))
p + 
  geom_polygon(data = world,
               aes(x=long, y=lat, group=group),
               color = "lightgrey", fill = "darkgrey") +
  geom_point(aes(color = temp, size = z, order = desc(z))) +
  # facet_wrap(~ months) +
  coord_map(xlim = xxlim, ylim = yylim)


```


###time plot
```{r, fig.height=3}
df3 <- subset(df2, date > as.Date("2011-08-03") & date < as.Date("2011-08-04"))
p <- ggplot(df3, aes(datetime, -z))
p + geom_point(aes(color = temp))

```


### using plot.ly

```{r, message=F, warning=F}
require(plotly)
plot_ly(df3, z = z, x = lon, y = lat, type = "scatter3d", mode = "markers", color = temp)


```


