---
title: "Practical training Rmarkdown"
author: "Willem Stolte"
date: "February 25, 2016"
output: pdf_document
---

## Start a new Rmarkdown document

* In Rstudio, choose from file menu: New File / R Markdown...
* Determine what kind of document you want it to be
    + for this training, choose "html"
    Note that you can also choose a presentation, which changes the choices of the formats that it can handle
* The document opens with some standard, functional, content to get going
* Note that the document contains both "normal" text (which actually is Markdown text) combined with 'chunks' of R code (the grey pieces)
* Try and run the document by pressing the "Knit html" button just above the document in Rstudio
    + an html document appears in which the Rmarkdown + R code is combined to a single document.
    
Add new data
Now, we are going to use data from an online source. We will use data from the Marine Projects database. 

Code can be inserted by "insert a new code chunk" button (green c) or by ctrl-alt-i 

Make a document that reads data, and uses the data for a time plot.

read data from the url: "http://marineprojects.openearth.nl/geoserver/mep-nsw/ows?service=WFS&version=1.0.0&request=GetFeature&typeName=mep-nsw:mep-nsw_bivalves&maxFeatures=50&outputFormat=csv"
```{r}
## define url of WFS service, where data are served as csv file:
url <- "http://marineprojects.openearth.nl/geoserver/mep-nsw/ows?service=WFS&version=1.0.0&request=GetFeature&typeName=mep-nsw:mep-nsw_bivalves&maxFeatures=50&outputFormat=csv"
## read data as csv (strings are read as character type, not as factors)
bivalves <- read.csv(url, stringsAsFactors = F)
```

Observe that these are only the first 50 records. This is fine for testing.

```{r}
##Look at the structure of the data set in the environment pane, or by running:
str(bivalves)
## Date and time are characters at this moment. They can be converted to date and time object by combining them in a new column 'datetime':
bivalves$datetime <- paste(bivalves$Begindatum, bivalves$Begintijd)
## and convert to POSIXct format
bivalves$datetime <- as.POSIXct(bivalves$datetime, format = "%Y-%m-%d %H:%M")
```

First, we plot locations, longitude and latitude fields are used.
Using Base functionality
```{r, fig.height = 3}
plot(bivalves$GeometriePunt.x, bivalves$GeometriePunt.y)
```

Or using the ggplot package

```{r PlotGgplot, fig.height=4}
require(ggplot2)
p <- ggplot(bivalves, aes(GeometriePunt.x, GeometriePunt.y))
p + geom_point(aes(color = Metingomschrijving))
```

Or using an interactive map

```{r PlotLeaflet}
require(leaflet)

leaflet() %>%
  addTiles(group = "OSM (default)") %>%
  addCircleMarkers(data=bivalves, 
                   lng = bivalves$GeometriePunt.x, 
                   lat = bivalves$GeometriePunt.y,
                   popup = sprintf(as.character(bivalves$Metingomschrijving)) 
                   # clusterOptions = markerClusterOptions()
                   )

```



